### 2013-06-20 doc/pgas13/flat-combining.tex
Some edits based on feedback from Brad.
A couple minor things based on Jacob's feedback.

\subsection{In Grappa}

@@

In distributed flat-combining, each core builds its own publication list to track all the operations waiting to be committed to the global data structure.
In Grappa, for each global data structure, a local \emph{proxy} object is allocated from the core-local heap to intercept requests.
Conceptually, workers add their requests to a local publication list in the proxy, one is chosen to do the combining, and the rest block until their request is satisfied.
However, because Grappa's scheduler is non-preemptive, each worker has atomicity ``for free'' until it performs a blocking operation (such as communication).
This means that an explicit publication list with clever synchronization is unnecessary.
Instead, workers merge their operations directly into the local proxy, and block if unable to satisfy their requests immediately.
The proxy structure is chosen to be able to compactly aggregate operations and efficiently perform matching in cases where it is allowed. Figure~\ref{fig:stackfc} shows how pushes and pops are matched on the proxy's local stack.

@@

Combining set and map operations expose more nuances in the consistency guarantees.
Insert and lookup operations performed by different tasks are inherently unordered unless synchronized externally.
Therefore, a batch of these operations can be committed globally in parallel, since they are guaranteed not to conflict with each other.
Note that if an insert finds that its key is already in the combiner, it does not need to send its own message. However, it must still block until that insert is done to ensure that a subsequent operation it performs cannot be reordered with it, preserving program order.
% providing the same guarantee as all other delegate operations. 
% This ensures that if one wanted to use the insert to signal synchronization, it is guaranteed to have been committed globally before the task continues.

Similarly, lookups can piggy-back on other lookups of the same key.
Using intuition from store/write buffers in modern processors, it is tempting to allow a lookup to be satisfied locally by inspecting outstanding inserts.
However, this allows the local order in which keys were inserted to be observed.
If this is allowed, then to preserve SC, that same order must be respected globally, forcing each batch to commit atomically with respect to other cores' batches. Enforcing this is prohibitively expensive, so a cheaper solution is chosen: if lookups must get their values from the global data structure, then batches can be performed in parallel.

@@

\section{Grappa FC Framework}
To leverage the flat-combining paradigm in Grappa, we implemented a generic framework to improve performance for a number of global data structures. The FC framework handles the common problems of managing the combiners, handling worker wake-ups, and maintaining progress. When hooking into the framework, each data structure implementation need only define how to combine operations and globally commit them.

>>>>

\subsection{Grappa}

@@

In distributed flat-combining, each core builds its own publication list to track all the operations waiting to be committed to the global data structure.
In Grappa, for each global data structure, a local \emph{proxy} object is allocated from the core-local heap to intercept requests.
Conceptually, workers add their requests to a local publication list in the proxy, one is chosen to do the combining, and the rest block until their request is satisfied.
However, because Grappa's scheduler is non-preemptive, each worker has atomicity ``for free'' until it performs a blocking operation (such as communication).
This means that an explicit publication list with clever synchronization is unnecessary.
Instead, workers merge their operations directly into the local proxy, and 
% block if unable to satisfy their requests immediately.
block, except in restricted cases where they are able to satisfy their requirements immediately without violating ordering rules for consistency (discussed in the next section).
The proxy structure is chosen to be able to compactly aggregate operations and efficiently perform matching in cases where it is allowed. Figure~\ref{fig:stackfc} shows how pushes and pops are matched on the proxy's local stack.

@@

Combining set and map operations expose more nuances in the consistency guarantees.
Insert and lookup operations performed by different tasks are inherently unordered unless synchronized externally.
Therefore, a batch of these operations can be committed globally in parallel, since they are guaranteed not to conflict with each other.
Note that if an insert finds that its key is already in the combiner object, it does not need to send its own message. However, it must still block until that insert is done to ensure that a subsequent operation it performs cannot be reordered with it, preserving program order.
% providing the same guarantee as all other delegate operations. 
% This ensures that if one wanted to use the insert to signal synchronization, it is guaranteed to have been committed globally before the task continues.

Similarly, lookups can piggy-back on other lookups of the same key.
Using intuition from store/write buffers in modern processors, it is tempting to allow a lookup to be satisfied locally by inspecting outstanding inserts.
However, this allows the local order in which keys were inserted to be observed.
If this is allowed, then to preserve SC, that same order must be respected globally, forcing each batch to commit atomically with respect to other cores' batches. Enforcing this is prohibitively expensive, so a cheaper solution is chosen: lookups must get their values from the global data structure, so batches can be performed in parallel.

@@

\section{Grappa FC Framework}
To leverage the flat-combining paradigm in Grappa, we implemented a generic framework to improve performance for a number of global data structures. The FC framework handles the common problems of managing the combiners, handling worker wake-ups, and maintaining progress. When hooking into the framework, each data structure need only define how to combine operations and globally commit them.

### 2013-06-20 doc/pgas13/intro.tex
A couple minor things based on Jacob's feedback.

% Globally shared data structures are one of the cornerstones of shared-memory and PGAS abstractions. 
% In order for multiple concurrent threads to interact with one another correctly, they must observe shared data consistently.
It is commonly accepted that the easiest consistency model to reason about is \emph{sequential consistency} (SC), which enforces that all accesses appear to happen in some global serializable order.
To preserve SC, operations on shared data structures should be \emph{linearizable}~\cite{herlihy1990linearizability}; that is, appear to happen atomically in some global total order.
In both physically shared memory and PGAS, maintaining linearizability presents performance challenges.
The simplest way is to have a single global lock to enforce atomicity and linearizability through simple mutual exclusion. Literally serializing accesses in this way is typically considered prohibitively expensive, even in physically shared memory.
% Much prior work has explored the scalability challenges of adding more concurrent accessors to shared data structures in shared memory systems.
However, as the number of concurrent accessors increases, the more they contend and the more synchronization operations fail, even for fine-grained lock-free synchronization schemes.
With the massive amount of parallelism in a cluster of multiprocessors and with the increased cost of remote synchronization, the problem is magnified.

>>>>

% Globally shared data structures are one of the cornerstones of shared-memory and PGAS abstractions. 
% In order for multiple concurrent threads to interact with one another correctly, they must observe shared data consistently.
It is commonly accepted that the easiest consistency model to reason about is \emph{sequential consistency} (SC), which enforces that all accesses are visible in program order and appear to happen in some global serializable order.
To preserve SC, operations on shared data structures should be \emph{linearizable}~\cite{herlihy1990linearizability}; that is, appear to happen atomically in some global total order.
In both physically shared memory and PGAS, maintaining linearizability presents performance challenges.
The simplest way is to have a single global lock to enforce atomicity and linearizability through simple mutual exclusion. Literally serializing accesses in this way is typically considered prohibitively expensive, even in physically shared memory.
% Much prior work has explored the scalability challenges of adding more concurrent accessors to shared data structures in shared memory systems.
However, as the number of concurrent accessors increases, the more they contend and the more synchronization operations fail, even for fine-grained lock-free synchronization schemes.
With the massive amount of parallelism in a cluster of multiprocessors and with the increased cost of remote synchronization, the problem is magnified.
